Geometric variational problems

### 2020-11-05 NSF2020/Description/specificaim1_ver2.tex
added bibs

The modern theory of membrane continuum mechanics
was pioneered by Hamm and Kozlov (HK) \cite{HaKo2000}.
The theory is widely used to describe biological phenomona 
fission \cite{10.1016/j.chemphyslip.2014.07.006, 10.1038/nature14509, 10.1103/PhysRevE.79.031926},
fusion \cite{10.1016/S0006-3495(02)75450-7,10.1073/pnas.121191898,10.1073/pnas.1119442109} ,
poration \cite{10.1016/j.bpj.2019.11.2221}, phase boundaries and interaction with inclusions
\cite{https://doi.org/10.1038/nrm.2017.16,10.1016/j.bpj.2019.11.2209, https://doi.org/10.1038/s41598-020-61110-2}, which require resolution of the internal structure of the membrane.  
%The Helfrich hamiltonian is widely used to determine energy barriers of fusion between lipid bilayer membranes \cite{ChKo08}. 
%Some of these barrier heights were never before obtained through the use of continuum membrane mechanics.
Using the HK  model as a base,
PI Ryham and collaborators calculated, for the first time in continuum theory, a least energy path for transitions between planar bilayers, a membrane stalk,
hemifusion diaphragm and the fusion pore \cite{RyWaCo13, RyKlYaCo16},
%The calculation of hydrophobic 
%fissures in \cite{RyKlYaCo16} motivated the HAP theory.
and the energies calculated by these continuum studies (Figure \ref{fig:barriers}A),
are in agreement with barrier heights derived by molecular dynamics and experimental studies \cite{FrRoPi17}.  
%analyzed the dependence of energy barriers on
%lipid composition and

@@

The twist deformation is a fully three-dimensional deformation. 
(Specific Aim 3 address outstanding implementation issues like three-dimensional boundary integral solvers.) 
The twist deformation is closely related to the assumption of lateral fluidity in membranes and so researchers 
have sometimes assumed $\KT = 0$ \cite{HaKo2000, 10.1063/1.4990404, https://doi.org/10.1039/C9SM02079A, 10.1103/PhysRevE.102.042406}, 
while molecular dynamics investigations find a twist modulus in the range of 1 to 2 \kBT \cite{LeVeWa14}.
To measure the elastic response,  we initially twist the bilayer along an $x_1$-coordinate axis.
In this configuration, the directors behave at leading order as $n = (0, mx_1, \sqrt{1 - m^2x_1^2})$,
yielding a surface gradient $\nabla n = \begin{pmatrix}0 & 0\\ m & 0\end{pmatrix}$. 
Here $\Div n = \mathrm{tr}(\nabla n) = 0$, $\det(\nabla n) = 0$ and $\Curl d = (\nabla u)_{12} - (\nabla u)_{21} = m$,
so that the continuum elastic energy equals $\tfrac{1}{2}\KT m^2 A + O(1)$ for a membrane area $A \gg 1$. (Our particle collections
conserve membrane area, unless they lie in a strong background flow.) 
This enables us to plot the HARP energy values derived from steepest gradient descent as a function of $m$, and finally
yields a value for $\KA$.

@@

 Perhaps surprisingly, the field of membrane continuum mechanics still lacks consensus as to
  whether \eqref{ansatz3} contains a complete list of consistent deformations.
  The PI Ryham showed, for example, that the tilt vector $T = n/n\cdot N - N$ leads to unphysical cusps when constraining membrane thickness
 \cite{RyKlYaCo16}.
  Recently, \cite{10.1063/1.4990404} derived a tilt curvature term that was neglected from the HK analysis \cite{HaKo2000}.
  Later, \cite{https://doi.org/10.1039/C9SM02079A} 
  and \cite{10.1103/PhysRevE.102.042406} independently identified an inconsistency \cite{10.1063/1.4990404} arising
  from a transversal tilt invariance assumption.

@@

  We will revisit the HK analysis to see if any inconsitencies can be resolved.
  As a main analytical innovation, we have found that expanding the strain tensor in
  terms of the coordinate frame $\{e'_1, e'_2, n\}$ with $e'_i$ perpendicular to $n$
  greatly simplifies the expansion because it decouples the gradient terms. As an illustration, the prominent works
  \cite{10.1063/1.4990404, 10.1103/PhysRevE.102.042406, HaKo2000, https://doi.org/10.1039/C9SM02079A} utilize an
  incompressility condition for the lipid elongation function $\zeta$ from \eqref{LMdeformation} that is only approximately true. 
  However, when expressed in the $\{e'_1, e'_2, n\}$ coordinate frame, the incompressility condition is exact
  and is given by a Steiner-type polynomial in terms of $\Div n$ and $\det \nabla n$ \cite{http://www.jstor.org/stable/1993504}.
  Aside from the special case of zero tilt, \cite{HaKo2000} an coworkers seem to have been unaware of this result.

>>>>

The modern theory of membrane continuum mechanics
was pioneered by Hamm and Kozlov (HK) \cite{Hamm2000}.
The theory is widely used to describe biological phenomona 
fission \cite{FrEsAkSh15, Maetal15, PhysRevE.79.031926},
fusion \cite{KoKo2002,10.1073/pnas.121191898,10.1073/pnas.1119442109} ,
poration \cite{10.1016/j.bpj.2019.11.2221}, phase boundaries and interaction with inclusions
\cite{https://doi.org/10.1038/nrm.2017.16,10.1016/j.bpj.2019.11.2209, https://doi.org/10.1038/s41598-020-61110-2}, which require resolution of the internal structure of the membrane.  
%The Helfrich hamiltonian is widely used to determine energy barriers of fusion between lipid bilayer membranes \cite{ChKo08}. 
%Some of these barrier heights were never before obtained through the use of continuum membrane mechanics.
Using the HK  model as a base,
PI Ryham and collaborators calculated, for the first time in continuum theory, a least energy path for transitions between planar bilayers, a membrane stalk,
hemifusion diaphragm and the fusion pore \cite{RyWaCo13, RyKlYaCo16},
%The calculation of hydrophobic 
%fissures in \cite{RyKlYaCo16} motivated the HAP theory.
and the energies calculated by these continuum studies (Figure \ref{fig:barriers}A),
are in agreement with barrier heights derived by molecular dynamics and experimental studies \cite{FrRoPi17}.  
%analyzed the dependence of energy barriers on
%lipid composition and

@@

The twist deformation is a fully three-dimensional deformation. 
(Specific Aim 3 address outstanding implementation issues like three-dimensional boundary integral solvers.) 
The twist deformation is closely related to the assumption of lateral fluidity in membranes and so researchers 
have sometimes assumed $\KT = 0$ \cite{Hamm2000, 10.1063/1.4990404, https://doi.org/10.1039/C9SM02079A, 10.1103/PhysRevE.102.042406}, 
while molecular dynamics investigations find a twist modulus in the range of 1 to 2 \kBT \cite{LeVeWa14}.
To measure the elastic response,  we initially twist the bilayer along an $x_1$-coordinate axis.
In this configuration, the directors behave at leading order as $n = (0, mx_1, \sqrt{1 - m^2x_1^2})$,
yielding a surface gradient $\nabla n = \begin{pmatrix}0 & 0\\ m & 0\end{pmatrix}$. 
Here $\Div n = \mathrm{tr}(\nabla n) = 0$, $\det(\nabla n) = 0$ and $\Curl d = (\nabla u)_{12} - (\nabla u)_{21} = m$,
so that the continuum elastic energy equals $\tfrac{1}{2}\KT m^2 A + O(1)$ for a membrane area $A \gg 1$. (Our particle collections
conserve membrane area, unless they lie in a strong background flow.) 
This enables us to plot the HARP energy values derived from steepest gradient descent as a function of $m$, and finally
yields a value for $\KA$.

@@

 Perhaps surprisingly, the field of membrane continuum mechanics still lacks consensus as to
  whether \eqref{ansatz3} contains a complete list of consistent deformations.
  The PI Ryham showed, for example, that the tilt vector $T = n/n\cdot N - N$ leads to unphysical cusps when constraining membrane thickness
 \cite{RyKlYaCo16}.
  Recently, \cite{10.1063/1.4990404} derived a tilt curvature term that was neglected from the HK analysis \cite{Hamm2000}.
  Later, \cite{https://doi.org/10.1039/C9SM02079A} 
  and \cite{10.1103/PhysRevE.102.042406} independently identified an inconsistency \cite{10.1063/1.4990404} arising
  from a transversal tilt invariance assumption.

@@

  We will revisit the HK analysis to see if any inconsitencies can be resolved.
  As a main analytical innovation, we have found that expanding the strain tensor in
  terms of the coordinate frame $\{e'_1, e'_2, n\}$ with $e'_i$ perpendicular to $n$
  greatly simplifies the expansion because it decouples the gradient terms. As an illustration, the prominent works
  \cite{10.1063/1.4990404, 10.1103/PhysRevE.102.042406, Hamm2000, https://doi.org/10.1039/C9SM02079A} utilize an
  incompressility condition for the lipid elongation function $\zeta$ from \eqref{LMdeformation} that is only approximately true. 
  However, when expressed in the $\{e'_1, e'_2, n\}$ coordinate frame, the incompressility condition is exact
  and is given by a Steiner-type polynomial in terms of $\Div n$ and $\det \nabla n$ \cite{http://www.jstor.org/stable/1993504}.
  Aside from the special case of zero tilt, \cite{Hamm2000} an coworkers seem to have been unaware of this result.
